Do we need detailed demographic data to forecast plant population responses to climate change?

### 2014-12-21 manuscript/manuscript.Rmd
updated ms formatting

---
layout: 12pt,review
linenumbers: true
title: "Do we need detailed demographic data to forecast plant population responses to climate change?"
author: 
  - name: Andrew T. Tredennick
    affiliation: usu
    email: atredenn(at)gmail.com
    footnote: Corresponding author
  - name: Peter B. Adler
    affiliation: usu
address: 
  - code: usu
    address: | 
      Department of Wildland Resources and the Ecology Center, 
      Utah State University,
      Logan, UT 84322, USA
abstract: |
  Forecasts of populations under future climate change are constrained by the spatial and temporal scales at which we can gather data. For plants, ecologists often require individual level data to build population models; data that is difficult and expensive to collect at spatial scales greater than one square meter. Models built using such data are difficult to scale-up to spatial scales relevant to land management decisions and policy intervention. A possible route forward are density-structured models based on coarse-scale census data, which are much easier to collect over large spatial extents. However, to use coarse-scale data in a climate change context requires the assumption that the climate signals in population dynamics identified using individual-level data can also be identified when individual-level data are averaged over. Here we use a longterm dataset to model population dynamics using individual-level and plot-level data.

keywords:
  - ecological forecasts
  - population models
  - integral projection model
  - climate change
  - grassland

bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl
documentclass: components/elsarticle

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    template: components/elsarticle.latex
    keep_tex: true

@@

@@

Introduction
============
Population models are important tools for predicting the impacts of environmental change on species. But reconciling the scales at which population models are parameterized and the scales at which environmental changes play out remains a challenge [@Freckleton2011; @Queenborough2011]. The major hurdle is that most population models, at least for plant species, are built using data from small, localized plots because parameterizing traditional population models requires tracking the fates of individuals. These models are difficult to scale up from the micro to meso-scales because the fitted parameters do not fully represent the spatial variation present at scales beyond that at which the data are collected [@Sæther2007]. Thus, our ability to use population models to predict the consequences of climate change is limited when we rely on individual-level data.

@@

Materials and Methods
=====================
### Study site and data
Our demographic data comes from the Fort Keogh Livestock and Range Research Laboratory in eastern Montana's northern mixed prairie near Miles City, Montana, USA (46 deg. 19' N, 105 deg 48' W). The dataset is freely available on Ecological Archives (CITE), and interested readers should refer to the metadata therin for a complete description. The site is about 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitaion falling from April through September (76). The site is grass dominated and, for the purposes of our study, we focus on the four most abundant graminoid species: \emph{Bouteloua gracilis}, \emph{Hesperostipa comata}, \emph{Pascopyrum smithii}, and \emph{Poa secunda}.

@@

Results
=======
We assessed the statistical importance of the climate covariates included the final vital rate regressions by comparing the residual deviance of models with climate covariates and temporal random effects, climate covariates only, and temporal random effects only. When a model includes climate covariates, this comparison shows the relative contribution of the climate covariates in explaining the total interannual variability [@Adler2012].

@@

References
==========

>>>>

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
fontsize: 12pt
geometry: margin=1in

@@

Do we need detailed demographic data to forecast population responses to climate change?
========================================================================================

### Andrew T. Tredennick and Peter B. Adler

*Andrew T. Tredennick ([atredenn@gmail.com](mailto:atredenn@gmail.com)), Department of Wildland Resources and the Ecology Center, Utah State University, Logan, UT*

*Peter B. Adler, Department of Wildland Resources and the Ecology Center, Utah State University, Logan, UT*

@@

Abstract
--------

Introduction
------------
Population models are important tools for predicting the impacts of environmental change on species. But reconciling the scales at which population models are parameterized and the scales at which environmental changes play out remains a challenge [@Freckleton2011; @Queenborough2011]. The major hurdle is that most population models, at least for plant species, are built using data from small, localized plots because parameterizing traditional population models requires tracking the fates of individuals. These models are difficult to scale up from the micro to meso-scales because the fitted parameters do not fully represent the spatial variation present at scales beyond that at which the data are collected [@Sæther2007]. Thus, our ability to use population models to predict the consequences of climate change is limited when we rely on individual-level data.

@@

Materials and Methods
---------------------
### Study site and data
Our demographic data comes from the Fort Keogh Livestock and Range Research Laboratory in eastern Montana's northern mixed prairie near Miles City, Montana, USA (46 deg. 19' N, 105 deg 48' W). The dataset is freely available on Ecological Archives [@Anderson2011], and interested readers should refer to the metadata therin for a complete description. The site is about 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitaion falling from April through September (76). The site is grass dominated and, for the purposes of our study, we focus on the four most abundant graminoid species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE).

@@

Results
-------
We assessed the statistical importance of the climate covariates included the final vital rate regressions by comparing the residual deviance of models with climate covariates and temporal random effects, climate covariates only, and temporal random effects only. When a model includes climate covariates, this comparison shows the relative contribution of the climate covariates in explaining the total interannual variability [@Adler2012].

@@

References
----------

### 2014-12-21 manuscript/manuscript.tex
updated ms formatting

\documentclass[author-year, 12pt,review]{components/elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add 
  \linenumbers % turns line numbering on 
\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{\it \hfill\today}%
 \let\@evenfoot\@oddfoot}
\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Do we need detailed demographic data to forecast plant population responses to climate change?},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{Do we need detailed demographic data to forecast plant population
responses to climate change?}
    \author[usu]{Andrew T. Tredennick\corref{c1}}
   \ead{atredenn(at)gmail.com} 
   \cortext[c1]{Corresponding author}
    \author[usu]{Peter B. Adler}

      \address[usu]{Department of Wildland Resources and the Ecology Center, Utah State
University, Logan, UT 84322, USA}

  \begin{abstract}
  Forecasts of populations under future climate change are constrained by
  the spatial and temporal scales at which we can gather data. For plants,
  ecologists often require individual level data to build population
  models; data that is difficult and expensive to collect at spatial
  scales greater than one square meter. Models built using such data are
  difficult to scale-up to spatial scales relevant to land management
  decisions and policy intervention. A possible route forward are
  density-structured models based on coarse-scale census data, which are
  much easier to collect over large spatial extents. However, to use
  coarse-scale data in a climate change context requires the assumption
  that the climate signals in population dynamics identified using
  individual-level data can also be identified when individual-level data
  are averaged over. Here we use a longterm dataset to model population
  dynamics using individual-level and plot-level data.
  \end{abstract}
   \begin{keyword} ecological forecasts \sep population models \sep integral projection model \sep climate change \sep grassland \sep \end{keyword}
 \end{frontmatter}

\section{Introduction}\label{introduction}

@@

\section{Materials and Methods}\label{materials-and-methods}

@@

Our demographic data comes from the Fort Keogh Livestock and Range
Research Laboratory in eastern Montana's northern mixed prairie near
Miles City, Montana, USA (46 deg. 19' N, 105 deg 48' W). The dataset is
freely available on Ecological Archives (CITE), and interested readers
should refer to the metadata therin for a complete description. The site
is about 800 m above sea level and mean annual precipitation (1878-2009)
is 334 mm, with most annual precipitaion falling from April through
September (76). The site is grass dominated and, for the purposes of our
study, we focus on the four most abundant graminoid species:
\emph{Bouteloua gracilis}, \emph{Hesperostipa comata},
\emph{Pascopyrum smithii}, and \emph{Poa secunda}.

@@

\section{Results}\label{results}

@@

\section{References}\label{references}

@@

>>>>

\documentclass[12pt,]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
% Redefine \includegraphics so that, unless explicit options are
% given, the image width will not exceed the width of the page.
% Images get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\ScaleIfNeeded{%
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother
\let\Oldincludegraphics\includegraphics
{%
 \catcode`\@=11\relax%
 \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
}%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\author{}
\date{}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing

\begin{document}

\normalsize

\section{Do we need detailed demographic data to forecast population
responses to climate
change?}\label{do-we-need-detailed-demographic-data-to-forecast-population-responses-to-climate-change}

\subsubsection{Andrew T. Tredennick and Peter B.
Adler}\label{andrew-t.-tredennick-and-peter-b.-adler}

\emph{Andrew T. Tredennick
(\href{mailto:atredenn@gmail.com}{\href{mailto:atredenn@gmail.com}{atredenn@gmail.com}}),
Department of Wildland Resources and the Ecology Center, Utah State
University, Logan, UT}

\emph{Peter B. Adler, Department of Wildland Resources and the Ecology
Center, Utah State University, Logan, UT}

\subsection{Abstract}\label{abstract}

\subsection{Introduction}\label{introduction}

@@

\subsection{Materials and Methods}\label{materials-and-methods}

@@

Our demographic data comes from the Fort Keogh Livestock and Range
Research Laboratory in eastern Montana's northern mixed prairie near
Miles City, Montana, USA (46 deg. 19' N, 105 deg 48' W). The dataset is
freely available on Ecological Archives (Anderson et al. 2011), and
interested readers should refer to the metadata therin for a complete
description. The site is about 800 m above sea level and mean annual
precipitation (1878-2009) is 334 mm, with most annual precipitaion
falling from April through September (76). The site is grass dominated
and, for the purposes of our study, we focus on the four most abundant
graminoid species: \emph{Bouteloua gracilis} (BOGR),
\emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and
\emph{Poa secunda} (POSE).

@@

\subsection{Results}\label{results}

@@

\subsection{References}\label{references}

@@

Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
92:1703.
